---
title: "Morton Analysis Report"
author: "Young Oh"
date: '2022-07-13'
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Working procedure

The Question we are interested in;

1) To what extent does root exudation rate vary among species? [ANOVA, though we're limited by study design - the 4-6 individual trees of each species are all located in the same plot, which isn't ideal for test assumptions.]
2) Does root exudation rate vary with plant or root traits? Which variables best predict root exudation rates? [ordination/PERMANOVA; mixed linear models that are similar to ANOVA and regressions but take plot-level replicates into account]
3) What are the relationships between root exudation rates (and other plant and root traits) and soil properties (rhizosphere effects in N cycling; aggregate stability)? [ordination/PERMANOVA; mixed linear models that are similar to ANOVA and regressions but take plot-level replicates into account]

## Before starting to analyze..

- The document will be update in Git hub (https://github.com/oh02rules/PhillipsLab) with the most updated code for the analysis
- Library list; more to come with multivariate analyses
library(plotrix)
library(ggpmisc)
library(ggplot2)
library(ggpubr)
library(magicaxis)
library(car)
library(emmeans)
library(pacman)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(ggeffects)
library(ggpmisc)

## Question 1) To what extent does root exudation rate vary among species?
To answer the question we perform an ANOVA on a fitted model with three factors, giving F-statistics
```{r}
setwd("/Users/young/Desktop/Git_Phillips_Lab/PhillipsLab/Morton_analyses")
rtrait <- read.csv("Exudation_Collection_compiled_072822.csv")
#generating the dataset with AM only
rtrait_AM <- rtrait[rtrait$Myco %in% c("AM"),]
#generating the dataset with EcM only
rtrait_EcM <- rtrait[rtrait$Myco %in% c("EM"),]

fit1 <- lm(Exudation_rate ~ Species, data=rtrait)#(Intercept) represent "B.Cypress"
fit2 <- lm(Exudation_rate ~ Myco, data=rtrait)
fit2_1 <- lm(Exudation_rate ~ Species + Myco, data=rtrait)
fit2_2 <- lm(Exudation_rate ~ Species, data=rtrait_AM)
fit2_3 <- lm(Exudation_rate ~ Species, data=rtrait_EcM)
fit3 <- lm(Exudation_rate ~ Phylo, data=rtrait)
fit4 <- lm(Exudation_rate ~ Leaf, data=rtrait)
summary(fit1)
summary(fit2)
summary(fit2_1)
summary(fit2_2)
summary(fit2_3)
summary(fit3)
summary(fit4)

boxplot(Exudation_rate~Species, data=rtrait)
par(mfrow=c(1,2))
boxplot(Exudation_rate~Species, data=rtrait_AM)
boxplot(Exudation_rate~Species, data=rtrait_EcM)

fit1_N <- lm(N_uptake ~ Species, data=rtrait) #(Intercept) represent "B.Cypress"
fit2_N <- lm(N_uptake ~ Myco, data=rtrait)
fit2_1_N <- lm(N_uptake ~ Species + Myco, data=rtrait)
fit2_2_N <- lm(N_uptake ~ Species, data=rtrait_AM)
fit2_3_N <- lm(N_uptake ~ Species, data=rtrait_EcM)#(Intercept) represent "Hemlock"
fit3_N <- lm(N_uptake ~ Phylo, data=rtrait)
fit4_N <- lm(N_uptake ~ Leaf, data=rtrait)
summary(fit1_N)
summary(fit2_N)
summary(fit2_1_N)
summary(fit2_2_N)
summary(fit2_3_N)
summary(fit3_N)
summary(fit4_N)

```
The ANOVA results show no statistical significant relationships in exudation rates among myco, Phylo and Myco types, while pawpaw show some significant difference in exudation rate due to their higher variability.

Next, to test whether the species were different from each other in exudation rate we use a multiple comparison test with emmeans function.
```{r}
# Load package
library(emmeans)
library(lme4)
fit1 <- lm(Exudation_rate ~ Species, data=rtrait)
# We estimate marginal means and confidence interval for each level of 'species'. 
tukey_test <- emmeans(fit1, 'Species')
pairs(tukey_test)

fit2 <- lm(Exudation_rate ~ Myco, data=rtrait)
tukey_test_Myco <- emmeans(fit2, 'Myco')
pairs(tukey_test_Myco)

fit1_N <- lm(N_uptake ~ Species, data=rtrait)
# We estimate marginal means and confidence interval for each level of 'species'. 
tukey_test_N <- emmeans(fit1_N, 'Species')
pairs(tukey_test_N)

fit2_N <- lm(N_uptake ~ Myco, data=rtrait)
tukey_test_Myco_N <- emmeans(fit2_N, 'Myco')
pairs(tukey_test_Myco_N)

```
The results of tukey contrasts show that exudation rates of all species (except Pawpaw) are not different from one another. 

Next, we perform ANOVA to compare nested models by performing a likelihood ratio test among models.
```{r}
#Testing whether species effects differ for different myco groups or are they simply additive?
fit_interaction <- lm(Exudation_rate ~ Myco + Species + Myco:Species, data=rtrait)
summary(fit_interaction)
fit3.2 <- lm(Exudation_rate ~ Myco * Species, data=rtrait)
fit3.3 <- lm(Exudation_rate ~ Species * Myco, data=rtrait)

library(car)
Anova(fit3.2)
Anova(fit3.3)
anova(fit1,fit2)
anova(fit1,fit3)
anova(fit1,fit4)
```
The ANOVA results testing a likelihood ratio on two nested models show all nested models are not significantly different.


## Figures related to ANOVA analyses

You can also embed plots, for example:
```{r, echo=FALSE, eval=FALSE}
library(ggpubr)
library(plotrix)
library(ggpmisc)
library(ggplot2)
library(ggpubr)
library(magicaxis)
library(car)
library(emmeans)
library(pacman)#This package is great for documents that you will be sharing – 
#it means that your collaborators will not have to spend time finding any packages they don’t have.
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(ggeffects)
```

```{r, echo=FALSE, eval=TRUE}
library(ggpubr)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(ggeffects)
library(ggpmisc)
AMvsECM <- rtrait %>% ggplot( aes(x = Myco, y = Exudation_rate, fill=Myco)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Boxplot for exudation rate among groups: Myco Type") +
  xlab("Myco. Type") + ylab(expression(log10~Exudation~rate~ ~(ugC~g^-1~day^-1))) + stat_compare_means(method = "anova")
AMvsECM + scale_y_continuous(trans='log10')

N_uptake_AMvsECM <- rtrait %>% ggplot( aes(x = Myco, y = N_uptake, fill=Myco)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Boxplot for Blank-corrected Net N uptake rate (mg N/(g root*day)) among groups: Myco Type") +
  xlab("Myco. Type") + ylab(expression(log10~Exudation~rate~ ~(mg~N/(g~root~day)))) + stat_compare_means(method = "anova")
N_uptake_AMvsECM

N_uptake_AMvsECM <- rtrait %>% ggplot( aes(x = Myco, y = N_uptake, fill=Myco)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Boxplot for Blank-corrected Net N uptake rate (mg N/(g root*day)) among groups: Myco Type") +
  xlab("Myco. Type") + ylab(expression(log10~Exudation~rate~ ~(mg~N/(g~root~day)))) + stat_compare_means(method = "anova")
N_uptake_AMvsECM

```

The ANOVA results show the nested models with respect to MF types are significantly different.

```{r, echo=FALSE, eval=TRUE}
#Boxplot for exudation rate among groups: Phylo types
Angio_vs_Gym <- rtrait %>% ggplot( aes(x = Phylo, y = Exudation_rate, fill=Phylo)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Boxplot for exudation rate among groups: Phylo Type") +
  xlab("Phylo Type") + ylab(expression(log10~Exudation~rate~ ~(ugC~g^-1~day^-1))) + stat_compare_means(method = "anova")
Angio_vs_Gym + scale_y_continuous(trans='log10')

N_uptake_Phylo <- rtrait %>% ggplot( aes(x = Phylo, y = N_uptake, fill=Phylo)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Blank-corrected Net N uptake rate (mg N/(g root*day)) among groups: Phylo Type") +
  xlab("Phylo. Type") + ylab(expression(Net~N~uptake~ ~(mg~N/(g~root~day)))) + stat_compare_means(method = "anova")
N_uptake_Phylo

N_uptake_Phylo_Myco <- rtrait %>% ggplot( aes(x = Phylo, y = N_uptake, fill=Myco)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Blank-corrected Net N uptake rate (mg N/(g root*day)) among groups: Phylo Type") +
  xlab("Phylo. Type with Myco type") + ylab(expression(Net~N~uptake~ ~(mg~N/(g~root~day)))) + stat_compare_means(method = "anova")
N_uptake_Phylo_Myco
```

The ANOVA results show the nested models with respect to Phylo. types (angiosperm vs. gymnosperm) are significantly different.

```{r, echo=FALSE, eval=TRUE}
#Boxplot for exudation rate among groups: Leaf types
dec_vs_ev <- rtrait %>% ggplot( aes(x = Leaf, y = Exudation_rate, fill=Leaf)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Boxplot for exudation rate among groups: Phylo Type") +
  xlab("Leaf Type") + ylab(expression(log10~Exudation~rate~ ~(ugC~g^-1~day^-1))) + stat_compare_means(method = "anova")
dec_vs_ev + scale_y_continuous(trans='log10')
#Not statistically different, 

N_uptake_Leaf <- rtrait %>% ggplot( aes(x = Leaf, y = N_uptake, fill=Leaf)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Boxplot for Blank-corrected Net N uptake rate (mg N/(g root*day)) among groups: Leaf Type") +
  xlab("Leaf. Type") + ylab(expression(Net~N~Uptake~rate~(mg~N/(g~root~day)))) + stat_compare_means(method = "anova")
N_uptake_Leaf

N_uptake_Leaf_Myco <- rtrait %>% ggplot( aes(x = Leaf, y = N_uptake, fill=Myco)) + geom_boxplot() +
  scale_fill_viridis(discrete = TRUE, alpha=0.6) +
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_ipsum() +
  theme(
    legend.position="none",
    plot.title = element_text(size=11)
  ) +
  ggtitle("Blank-corrected Net N uptake rate (mg N/(g root*day)) among groups: Leaf Type with Myco") +
  xlab("Leaf. Type") + ylab(expression(Net~N~Uptake~rate~(mg~N/(g~root~day)))) + stat_compare_means(method = "anova")
N_uptake_Leaf_Myco
```

The ANOVA results show the nested models with respect to leaf types (deciduous vs. evergreen) are not significantly different.


```{r, echo=FALSE, eval=TRUE}
library(ggpubr)
p1 <- ggplot(rtrait, aes(x=Species, y=Exudation_rate, fill=Myco)) + 
  geom_boxplot() +
  facet_wrap(~Myco) + geom_jitter(color="black", size=0.4, alpha=0.9) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(log10~Exudation~rate~ ~(ugC~g^-1~day^-1)))+
  ggtitle("Boxplot for exudation rate among groups: Myco. Type")

p1 + facet_wrap(~Myco) + scale_y_continuous(trans='log10')

p2 <- ggplot(rtrait, aes(x=Species, y=N_uptake, fill=Myco)) + 
  geom_boxplot() +
  facet_wrap(~Myco) + geom_jitter(color="black", size=0.4, alpha=0.9) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(Net~N~Uptake~rate~(mg~N/(g~root~day))))+
  ggtitle("Blank-corrected Net N uptake rate (mg N/(g root*day)): Myco. Type")

p2 + facet_wrap(~Myco)
```

The ANOVA results show the extent of statistical significance among species differs among different mycorrhizal associations showing EcM tree species vary more. The figure needs to be enlarged to show all the label on the y-axis.

## 2) Does root exudation rate vary with plant or root traits? Which variables best predict root exudation rates?

To answer the questions, we need to construct mixed linear models that are similar to ANOVA and regressions but take plot-level replicates into account. To be continued.

First, we fit the model with exudation rate as the response and Surf_area as the predictor and have a look at the output:
As a example, we relate surface area to exudation rate that were thought to have a causal relationship. 
```{r}
#Testing a plain corr. plot
prelim_plot <- ggplot(rtrait, aes(x = Surf_area, y = Exudation_rate)) +
geom_point(aes(color=Species)) + xlab(expression(Surface~Area~(cm^2))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1))) + 
geom_smooth(method = "lm") +  stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05)
prelim_plot

gd <- rtrait %>%  group_by(Species) %>% summarise(Surf_area = mean(Surf_area),
                  Exudation_rate  = mean(Exudation_rate))
gd

plot_RTD_spp_mean <- ggplot(rtrait, aes(x = Surf_area, y = Exudation_rate)) +
geom_point(alpha = .5, aes(color=Species, shape=Myco)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd, size = 4, aes(color=Species)) +  theme_bw() + guides(color = guide_legend("Species"),  shape = guide_legend("Species")) +
  labs(title = "Relationship of root exudation with Surf_area") + xlab(expression(Surf_area~(cm^2))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))
#geom_smooth(method = "lm")
plot_RTD_spp_mean + geom_smooth(method = "lm")



```

It appears to have some correlation here. But, we need to decompose effects to construct the best fitting model.

Further, we separate the correlation between surface area and exudation rate into different mycorrhizal types (AM vs EcM) since we found the significant difference in exudation rates through ANOVA. 
```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = Surf_area, y = Exudation_rate, shape=Myco)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Myco)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +labs(title = "Relationship of root exudation with Surf_area") + xlab(expression(Surface~area~(cm^2))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~Myco)

#AM only plot

plt_AM_only <- ggplot(rtrait_AM,aes(x = Surf_area, y = Exudation_rate)) + 
  geom_point(size=3, aes(shape=Species, color=Species)) +
  geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) + labs(title = "Relationship of root exudation with Surf_area among EM Myco type") + xlab(expression(Surface~area~(cm^2))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt_AM_only +scale_shape_manual(values = c(15, 16, 17, 18, 19, 20))

plt_EM_only <- ggplot(rtrait_EcM,aes(x = Surf_area, y = Exudation_rate)) + 
  geom_point(size=3, aes(shape=Species, color=Species)) + geom_smooth(method = "lm") + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) + labs(title = "Relationship of root exudation with Surf_area among EM Myco type") + xlab(expression(Surface~area~(cm^2))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt_EM_only + scale_shape_manual(values = c(15, 16, 17, 18, 19, 20 ))


```

It appears that the correlation between surface area and exudation rate is stronger in EcM species than AM species. We need to know why.

Since we already found out the correlation among root traits including exudation may differ among myco. type, we construct scatterplot Matrix. 

First, let's examine the correlation between root exudation and all root traits.

```{r, echo=FALSE}
library(car)
#spm(~Exudation_rate+Diam+SRL+SRA+RTD+Mass+Surf_area|Myco, data=rtrait, main="Scatterplot Matrix among root traits")

# another way to construct a scatterplot matrix with R sq. values

library(GGally)
#removing SampleID column
rtrait_new <- subset (rtrait, select = -c(SampleID,Myco,Phylo,Leaf))

ggpairs(rtrait_new)
```
Making a heat map

```{r eval=FALSE, include=FALSE}
#removing categorical variable column
rtrait_cor <- subset (rtrait, select = -c(SampleID,Species,Myco,Phylo,Leaf))

rtrait.cor = cor(rtrait_cor, method = c("spearman"))

install.packages("Hmisc")
library("Hmisc")
rtrait.rcorr = rcorr(as.matrix(rtrait_cor))
rtrait.rcorr
#This generates one table of correlation coefficients (the correlation matrix) and another table of the p-values.

#extract the values from this object into a useable data structure
rtrait.coeff = rtrait.rcorr$r
rtrait.p = rtrait.rcorr$P

cor_1 <- cor(rtrait_cor[1:10], method = c("spearman"), use = "complete.obs") 
library(dplyr)
symnum(cor_1, abbr.colnames = FALSE)
test_trait = cor.mtest(rtrait_cor, conf.level = 0.95)

corrplot.mixed(cor_1, order = 'AOE')

corrplot(cor_1, p.mat = test_trait$p, method = 'circle', type = 'lower', sig.level = c(0.001, 0.01, 0.05), pch.cex = 0.9, insig = 'label_sig', pch.col = 'grey20', addCoef.col ='black', number.cex = 0.8, order = 'AOE', diag=FALSE)

install.packages("corrplot")
library(corrplot)

#generating a correlation matrix plot (called a Correlogram); Positive correlations are displayed in a blue scale while negative correlations are displayed in a red scale.
corrplot(rtrait.cor)

#Another way to visualize spearman's correlation coefficients
install.packages("PerformanceAnalytics")
library("PerformanceAnalytics")

corrplot_spearman <- rtrait_cor[, c(1,2,3,4,5,6,7,8,9)]
chart.Correlation(rtrait_cor, histogram=TRUE, pch=19)

# double-checking corr. between exudation and RTD
cor.test(rtrait$Exudation_rate, rtrait$RTD, method = "spearman")

library(corrplot)

corrplot(rtrait.rcorr, type = "upper", order = "hclust", 
         tl.col = "black", tl.srt = 45)
```



```{r}
# More visualization with bivariate analysis

install.packages("GGally", repos = "http://cran.us.r-project.org")
library(GGally)
library(ggplot2)
library(Hmisc)
ggpairs(rtrait, columns = c("Exudation_rate", "Diam", "SRL", "SRA", "RTD","Surf_area", "Root_vol", "N_uptake" ), title = "Bivariate analysis of Exudation rates vs. root traits", upper = list(continuous = wrap("cor",
        size = 3)),
    lower = list(continuous = wrap("smooth",
        alpha = 0.3,
        size = 0.1)),
    mapping = aes(color = Myco))

```

The matrix shows that exudation rates in these samples are positively correlated with Diam, RTD, and SRA, while the degree of correlation differ among myco types. Especially, exudation rate of EM species can be explained by greater number of root morphological traits such as Diam, RTD, Surf_area, Root_vol, and SA of tips. We need to investigate if exudation rates of EM species are really more correlated with root morphology and why.   

Second, we separate AM and EcM in examining the correlation between root exudation and all root traits.

For AM,
```{r}
library(ggpmisc)
library(ggplot2)
#generating the dataset with AM only
rtrait_AM <- rtrait[rtrait$Myco %in% c("AM"),]
#generating the dataset with EcM only
rtrait_EcM <- rtrait[rtrait$Myco %in% c("EM"),]
#spm(~Exudation_rate+Diam+SRL+SRA+RTD+Mass+Surf_area|Myco, data=rtrait_AM, main="Scatterplot Matrix among root traits: AM trees")
#removing SampleID column
rtrait_AM_new <- subset (rtrait_AM, select = -c(SampleID,Myco,Diam))

ggpairs(rtrait_AM_new)

```

For EcM, 
```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)
library(ggplot2)
#spm(~Exudation_rate+Diam+SRL+SRA+RTD+Mass+Surf_area|Myco, data=rtrait_EcM, main="Scatterplot Matrix among root traits: EcM trees")
rtrait_EM_new <- subset (rtrait_EcM, select = -c(SampleID,Myco,Phylo,Leaf,Tip_Mass))

ggpairs(rtrait_EM_new)

```

## In-depth correlation analyses

Following the procedure we built to investigate the correlation between Surf_area and exudation rate, we examine a few root traits that appear to correlate with root exudation. 

Exudation rate vs. RTD

```{r, echo=FALSE, eval=TRUE}
prelim_plot_RTD <- ggplot(rtrait, aes(x = RTD, y = Exudation_rate)) +
geom_point(aes(color=Species)) +
geom_smooth(method = "lm") +  stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05)
prelim_plot_RTD

#adding group means for each species to the plot

#taking group mean

gd <- rtrait %>%  group_by(Species) %>% summarise(RTD = mean(RTD),
                  Exudation_rate  = mean(Exudation_rate))
gd

plot_RTD_spp_mean <- ggplot(rtrait, aes(x = RTD, y = Exudation_rate)) +
geom_point(alpha = .5, aes(color=Species, shape=Leaf)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd, size = 4, aes(color=Species)) +  theme_bw() + guides(color = guide_legend("Species"),  shape = guide_legend("Species")) + labs(title = "Relationship of root exudation with RTD") + xlab(expression(RTD~(g~cm^-3))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))
#geom_smooth(method = "lm")
plot_RTD_spp_mean + geom_smooth(method = "lm")





```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = RTD, y = Exudation_rate)) + 
  geom_point(alpha = .75, aes(color=Species, shape=Leaf)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +  theme_bw() + guides(color = guide_legend("Species"),  shape = guide_legend("Species")) + labs(title = "Relationship of root exudation with RTD (AM vs. EcM)") + xlab(expression(RTD~(g~cm^-3))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))
#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~Myco, scales = "free") + geom_smooth(method = "lm") 
```

Exudation rate vs. SRA

```{r, echo=FALSE, eval=TRUE}
prelim_plot_RTD <- ggplot(rtrait, aes(x = SRA, y = Exudation_rate)) +
geom_point(aes(color=Species, shape=Myco)) +
geom_smooth(method = "lm") +  stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) + labs(title = "Relationship of root exudation with SRA") + xlab(expression(SRA~(cm^2~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))
prelim_plot_RTD
```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = SRA, y = Exudation_rate, color=Myco, shape=Myco)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Myco)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 3.5, label.x.npc = "right", vstep = 0.05) + labs(title = "Relationship of root exudation with SRA (Myco Type: AM vs. EcM)") + xlab(expression(SRA~(cm^2~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

plt + facet_wrap(~Myco, scales = "free") + geom_smooth(method = "lm") 
```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = SRA, y = Exudation_rate, color=Leaf, shape=Leaf)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Leaf)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 3.5, label.x.npc = "right", vstep = 0.05)+ labs(title = "Relationship of root exudation with SRA (Leaf type: dec vs. ev)") + xlab(expression(SRA~(cm^2~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

plt + facet_wrap(~Leaf, scales = "free") + geom_smooth(method = "lm") 
```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = SRA, y = Exudation_rate, color=Phylo, shape=Phylo)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 3.5, label.x.npc = "right", vstep = 0.05)+ labs(title = "Relationship of root exudation with SRA (Angio vs. Gym)") + xlab(expression(SRA~(cm^2~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))


plt + facet_wrap(~Phylo, scales = "free") + geom_smooth(method = "lm") 
```

Exudation rate vs. SRL
```{r, echo=FALSE, eval=TRUE}
prelim_plot_SRL <- ggplot(rtrait, aes(x = SRL, y = Exudation_rate)) +
geom_point(aes(color=Myco)) +
geom_smooth(method = "lm") +  stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) + labs(title = "Relationship of root exudation with SRL") + xlab(expression(SRL~(cm~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))
prelim_plot_SRL
```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = SRL, y = Exudation_rate, color=Myco, shape=Myco)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Myco)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 3.5, label.x.npc = "right", vstep = 0.05)+ labs(title = "Relationship of root exudation with SRL (AM vs. EcM)") + xlab(expression(SRL~(cm~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

plt + facet_wrap(~Myco, scales = "free") + geom_smooth(method = "lm") 
```
```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = SRL, y = Exudation_rate, color=Phylo, shape=Phylo)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 3.5, label.x.npc = "right", vstep = 0.05)+ labs(title = "Relationship of root exudation with SRL (Angio vs. Gym)") + xlab(expression(SRL~(cm~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

plt + facet_wrap(~Phylo, scales = "free") + geom_smooth(method = "lm") 
```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = SRL, y = Exudation_rate, color=Leaf, shape=Leaf)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Leaf)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", size = 3.5, vstep = 0.05)+ labs(title = "Relationship of root exudation with SRL (dec vs. ev)") + xlab(expression(SRL~(cm~g^-1))) + ylab(expression(Exudation~rate~ ~(ugC~g^-1~day^-1)))

plt + facet_wrap(~Leaf)
```
Exudation rate vs. N uptake
```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)
prelim_plot_N <- ggplot(rtrait, aes(x = Exudation_rate, y = N_uptake)) +
geom_point(aes(color=Species, shape=Myco)) +
geom_smooth(method = "lm") +  stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) + labs(title = "Relationship of N uptake with exudation") + xlab(expression(Exudation~rate~(ugC~g^-1~day^-1))) + ylab(expression(Net~N~uptake~rate~(mg~ N~g^-1~day^-1)))
prelim_plot_N
```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = Exudation_rate, y = N_uptake, color=Myco, shape=Myco)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Myco)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.5, label.x.npc = "right", vstep = 0.05)+ labs(title = "Relationship of N uptake rate with root exudation (AM vs. EcM)") + xlab(Exudation~rate~ ~(ugC~g^-1~day^-1)) + ylab(expression(Net~N~uptake~rate~(mg~ N~g^-1~day^-1)))

plt + facet_wrap(~Myco, scales = "free") + geom_smooth(method = "lm") 
```
```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = Exudation_rate, y = N_uptake, color=Phylo, shape=Phylo)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 3.5, label.x.npc = "right", vstep = 0.05)+ labs(title = "Relationship of N uptake rate with root exudation (Angio vs. Gymno)") + xlab(Exudation~rate~ ~(ugC~g^-1~day^-1)) + ylab(expression(Net~N~uptake~rate~(mg~ N~g^-1~day^-1)))

plt + facet_wrap(~Phylo, scales = "free") + geom_smooth(method = "lm") 
```

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)
library(ggpmisc)

plt <- ggplot(rtrait,aes(x = Exudation_rate, y = N_uptake, color=Leaf, shape=Leaf)) + 
  geom_point(size=3, aes(color=Species)) + 
  geom_smooth(method = "lm", aes(fill=Leaf)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 3.5, label.x.npc = "right", vstep = 0.05)+ labs(title = "Relationship of N uptake rate with root exudation (dec vs. ev)") + xlab(Exudation~rate~ ~(ugC~g^-1~day^-1)) + ylab(expression(Net~N~uptake~rate~(mg~ N~g^-1~day^-1)))

plt + facet_wrap(~Leaf, scales = "free") + geom_smooth(method = "lm") 
```


Root traits related with root length appear to have no significant relationship with exudation. We need to know why.

##Linear mixed effect model

Because we are interested in which root traits are the best predictors to explain the variation in root exudation among different tree species and between functional types, here we are generating procedure for linear mixed effect model to evaluate the extent of fixed and random effects.

First, #First, we look at sphericity, which allows us to test variances of the differences in the pairs of repeated measured to be equal. Here, we test if epsilon=1, getting the range if the estimates are in between Greenhouse-Geisser method (lower bound) and Huynh-Feldt method (upper); the narrower the interval is, the better the estimates fit the assumption.

```{r, echo=FALSE, eval=TRUE}
library(nlme)
library(emmeans)
library(reshape)
library(ez)
library(psych)
library(ggplot2)
setwd("/Users/young/Desktop/Git_Phillips_Lab/PhillipsLab/Morton_analyses")
rtrait <- read.csv("Exudation_Collection_compiled_072822.csv")
# we may use the data that we separated AM and EcM 


```

Principal Components Analysis (PCA)

PCA is ultimately the transformation of continuous multivariate data into new variables that maximise the amount of 
variance explained in the data, with each subsequent variable orthogonal
```{r, echo=FALSE, eval=TRUE}
PCA_rtrait <- read.csv("Exudation_Collection_compiled_072822.csv")

head(PCA_rtrait)

rtrait_new <- subset (PCA_rtrait, select = -c(SampleID,Myco,Phylo,Leaf,Tip_Mass))
rda_PCA_rtrait <- rda(rtrait_new[,-1], scale=TRUE)
plot(rda_PCA_rtrait)
biplot(rda_PCA_rtrait, type = c("text", "points"))
summary(rda_PCA_rtrait, display=NULL)
screeplot(rda_PCA_rtrait, main=NULL)
screeplot(rda_PCA_rtrait, main=NULL, bstick=T, type='l')

#An example of cluster analysis 
plot(hclust(dist(scale(rtrait[,-c(1:5)]))))

#An example of PerMANOVA (Permutational ANOVA) analysis 
adonis(vegdist(rtrait[,6:14]) ~ Myco*Species, data=rtrait)

res.man <- manova(cbind(Exudation_rate, RTD) ~ Species, data = rtrait)
summary(res.man)

```

##Linear mixed model procedure

An exercise (framework) to build a linear mixed model that fixed and random effects were separated. Here, we consider Species as a random effect and plant groups (Myco, Phylo, and Leaf) as fixed effects.

We start with a box plot descirbing exudation rate differing between myco. types.

```{r, echo=FALSE, eval=TRUE}
library(ggplot2)
library(dplyr)
rtrait %>% ggplot(aes(x= Myco, y=Exudation_rate)) + geom_boxplot()
rtrait %>% ggplot(aes(x= Species, y=Exudation_rate)) +
  geom_boxplot() + facet_grid(.~Myco)
rtrait %>% ggplot(aes(x= Species, y=Exudation_rate)) +
  geom_jitter() +
  facet_grid(Myco~Species)
```

Syntax for linear mixed model
model = lmer(target ~ fixed1 + fixed2+  (1|random1) +  (1|random2), data=data)
```{r, echo=FALSE, eval=TRUE}
lm_model = lmer(Exudation_rate ~ Myco + (1|Species), data=rtrait)
lm_model
```
```{r, echo=FALSE, eval=TRUE}
lm_model2 = lmer(Exudation_rate ~ Myco + Phylo +(1|Species), data=rtrait)
lm_model2
```

```{r, echo=FALSE, eval=TRUE}
lm_model_null = lmer(Exudation_rate ~ Phylo +(1|Species), data=rtrait)
lm_model_null
```

```{r}
anova(lm_model2,lm_model_null)
```
Including Myco increases the robustness of prediction. Again, with Leaf type.

```{r, echo=FALSE, eval=TRUE}
lm_model3 = lmer(Exudation_rate ~ Myco + Leaf +(1|Species), data=rtrait)
lm_model3
```

```{r, echo=FALSE, eval=TRUE}
lm_model_null2 = lmer(Exudation_rate ~ Leaf +(1|Species), data=rtrait)
lm_model_null2
```

```{r}
anova(lm_model3,lm_model_null2)
```
Likewise, including Myco increases the robustness of prediction.



