---
title: "Morton_analyses"
author: "Young Oh"
date: '2022-09-09'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Working procedure

The Question we are interested in;

1) To what extent does root exudation rate vary among species? [ANOVA, though we're limited by study design - the 4-6 individual trees of each species are all located in the same plot, which isn't ideal for test assumptions.]
2) Does root exudation rate vary with plant or root traits? Which variables best predict root exudation rates? [ordination/PERMANOVA; mixed linear models that are similar to ANOVA and regressions but take plot-level replicates into account]
3) What are the relationships between root exudation rates (and other plant and root traits) and soil properties (rhizosphere effects in N cycling; aggregate stability)? [ordination/PERMANOVA; mixed linear models that are similar to ANOVA and regressions but take plot-level replicates into account]

## Before starting to analyze..

- The document will be update in Git hub (https://github.com/oh02rules/PhillipsLab) with the most updated code for the analysis
- Library list; more to come with multivariate analyses
```{r echo=FALSE,eval=FALSE}
library(plotrix)
library(dplyr)
library(ggpmisc)
library(ggplot2)
library(ggpubr)
library(magicaxis)
library(car)
library(emmeans)
library(pacman)
library(tidyverse)
library(hrbrthemes)
library(viridis)
library(ggeffects)
library(corrplot)
# do this if needed, install.packages("lmerTest", repos = "http://cran.us.r-project.org")
library(lmerTest)
library(ggpmisc)
library(nortest)

install.packages("lattice")                                  # Install lattice package
library("lattice")  
```
- Importing and creating the env. 
   1. Fixing up data
   2. Characterizing the sample distribution
```{r echo=FALSE,eval=FALSE}
setwd("/Users/young/Desktop/Git_Phillips_Lab/PhillipsLab/Morton_analyses/")
exudate <- read.csv("final exudate data_master.csv")
#generating the dataset with AM only
exudate_AM <- exudate[exudate$myco %in% c("am"),]
#generating the dataset with EcM only
exudate_EcM <- exudate[exudate$myco %in% c("em"),]

#change variable names to make easier
exudate$total_exudate <- exudate$total_exudate_μg_day
exudate$mass <- exudate$mass_g
exudate$tip_mass <- exudate$tip_mass_g
exudate$length <- exudate$length_cm
exudate$er_mass_tip <- as.numeric(exudate$er_mass_tip)
exudate$n <- exudate$X.N
exudate$c <- exudate$X.C
exudate$c_n <- exudate$C.N

#remove negative values and two inncaurate pawpaw observations
exudate <- exudate%>%
  filter(total_exudate > 1, er_mass < 11000)
```
First, the result of the Anderson-Darling test for normality show the exudation measurements are not normally distributed, calling for log-transformation. 

```{r echo=FALSE,eval=TRUE}
exudate <- read.csv("final exudate data_master.csv")
library(nortest)
qqnorm(exudate$total_exudate)
qqline(exudate$total_exudate)
ad.test(exudate$total_exudate)
#The result shows it's not normal, more or less rightly-skewed. Thus, we test if log-transformation is necessary
qqnorm(log(exudate$total_exudate))
qqline(log(exudate$total_exudate))
ad.test(log(exudate$total_exudate))

# log transform exudation values to make the data follow the normal distribution
exudate$er_mass_log <- log(exudate$er_mass)
exudate$er_length_log <- log(exudate$er_length)
exudate$er_sa_fine_log <- log(exudate$er_sa_fine)
exudate$er_sa_log <- log(exudate$er_sa)
exudate$er_mass_tip_log <- log(exudate$er_mass_tip)
exudate$total_exudate_log <- log(exudate$total_exudate) 



```
The second fig. shows log-transformation of the exudation data made distribution normal, allowing us to use best-available stat. methods (p-value of 0.90, meaning log-transformed data is now following the normal distribution. 

# To what extent does root exudation rate vary among species?

Strategy) We use ANOVA, though we're limited by study design in that the 4-6 individual trees of each species are all located in the same plot, which isn't ideal for test assumptions.

## Species differences in exudation rate and net N uptake rate

**Exudation rate (mass- vs area-specific)**

```{r, echo=FALSE, eval=TRUE}
library(ggplot2)
library(emmeans)
library(ggpubr)
p0 <- ggplot(exudate, aes(x=reorder(species, er_sa_fine_log), y=er_sa_fine_log, fill=species)) + 
  geom_boxplot() + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~cm^-2)))+
  ggtitle("Boxplot for exudation rate per fineroot area for 11 species")+ theme(axis.text.x = element_text(size = 10, angle= 45))
p0
p1 <- ggplot(exudate, aes(x=reorder(species, er_mass_log), y=er_mass_log, fill=species)) + 
  geom_boxplot() + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~g^-1)))+
  ggtitle("Boxplot for exudation rate per fineroot mass for 11 species")+ theme(axis.text.x = element_text(size = 10, angle= 45))
p1
```

No significant difference among species overall in both area- and mass-specific exudation, while there are differences between area- and mass-specific exudation rate, meaning there are some difference in efficiency of root upon exudation rate (e.g., RTD vs exudation).  

**Net N uptake rate**

```{r, echo=FALSE, eval=TRUE}
p5 <- ggplot(exudate, aes(x=reorder(species, er_mass_log), y=N_uptake, fill=species)) + 
  geom_boxplot() + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species") + ylab(expression(Net~N~uptake~rate~(mgN~root~g^-1~d^-1)))+
  ggtitle("Net N uptake rate for 11 species")+ theme(axis.text.x = element_text(size = 10, angle= 45))
p5


```

Although the results of all the species (n=11) show no significant difference in both exudation and net N uptake rate, there is a significant difference within EcM trees. The contrast in species difference between exudation rate and net N uptake rate is interesting in that the variability of net N uptake rates is smaller than that of exudation rates.

```{r, echo=FALSE, eval=FALSE}
#not for this analysis
# to see if we can plot two diff. exudation rates side by side for each species -> not working well
#reshaping exudate data

library(MASS) 
library(reshape2) 
library(reshape) 


reshape<- exudate[,c("species", "er_mass", "er_sa")]
head(reshape)
reshape_1<-melt(reshape, id = "species")
head(reshape_1)
p2 <- ggplot(reshape_1, aes(x = variable, y = value, color = species)) +  # ggplot function
  geom_boxplot()
p2 + scale_y_continuous(trans='log')
```

## Representation of species differences nested on plant functional groups (Myco, Phylo, and Leaf)

**Mycorrhizal types (AM vs. EcM)**

When exudation rates and N uptake rates for species are nested on myco. type, the figures show interesting results; exudation rates are significantly different among EM associated species, whereas N uptake are significantly different among AM associated species. We need to know why.  

```{r, echo=FALSE, eval=TRUE}
#generating the dataset with AM only
exudate_AM <- exudate[exudate$myco %in% c("am"),]
#generating the dataset with EcM only
exudate_EcM <- exudate[exudate$myco %in% c("em"),]
exudate_AM$er_mass_log <- log(exudate_AM$er_mass)
exudate_AM$er_sa_fine_log <- log(exudate_AM$er_sa_fine)
exudate_EcM$er_mass_log <- log(exudate_EcM$er_mass)

p8 <- ggplot(exudate_AM, aes(x=reorder(species, er_sa_fine_log), y=er_sa_fine_log, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~myco) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~cm^-2)))+
  ggtitle("Area-specific exudation rate among groups: AM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=0))

p8 + facet_wrap(~myco)

p9 <- ggplot(exudate_EcM, aes(x=reorder(species, er_sa_fine_log), y=er_sa_fine_log, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~myco) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~cm^-2)))+
  ggtitle("Area-specific exudation rate among groups: EcM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=0))

p9 + facet_wrap(~myco)

p10 <- ggplot(exudate_AM, aes(x=reorder(species, er_mass_log), y=er_mass_log, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~myco) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~g^-1)))+
  ggtitle("Mass-specific exudation rate among groups: AM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=0))

p10 + facet_wrap(~myco)

p11 <- ggplot(exudate_EcM, aes(x=reorder(species, er_mass_log), y=er_mass_log, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~myco) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~g^-1)))+
  ggtitle("Mass-specific exudation rate among groups: EcM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=0))

p11 + facet_wrap(~myco)
```

When separating exudation rates of the species into myco. types, we can observe interesting results showing EcM having significant differences while those of AM trees are not. Especially among EcM trees, as seen in lower exudation rates in hickory and SWO having broadleaf and angiosperm functional type, the differences in plant functional types may play a important role in their lower exudation rates. Brunn et al. (2022) reported no significant differences in area-specific between Fagus sylvatica and Picea abies, while a fraction of photosynthates allocated to exudation increases upon drought treatment for both species. This allows us to ponder whether the current edaphic conditions that vary specific to each monoculture plot may influence exudation rates and corresponding N uptake differences. 

```{r, echo=FALSE, eval=TRUE}
#generating the dataset with AM only
exudate_AM <- exudate[exudate$myco %in% c("am"),]
#generating the dataset with EcM only
exudate_EcM <- exudate[exudate$myco %in% c("em"),]
exudate_AM$er_mass_log <- log(exudate_AM$er_mass)
exudate_EcM$er_mass_log <- log(exudate_EcM$er_mass)

par(mfrow=c(1,2))
p6 <- ggplot(exudate_AM, aes(x=reorder(species, er_mass_log), y=N_uptake, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~myco) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(Net~N~uptake~rate~(mgN~root~g^-1~d^-1)))+
  ggtitle("Net N uptake rate among groups: AM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=90))
p6

p7 <- ggplot(exudate_EcM, aes(x=reorder(species, er_mass_log), y=N_uptake, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~myco) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(Net~N~uptake~rate~(mgN~root~g^-1~d^-1)))+
  ggtitle("Net N uptake rate among groups: EcM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=90))

p7

```

Similarly, EcM trees exhibit significant differences in net N uptake rates, while AM trees does not. 

# How did exudation & N uptake differ between myco type, phylo, and leaf habit?

```{r, echo=FALSE, eval=TRUE}
#generating the dataset with AM only
exudate_AM <- exudate[exudate$myco %in% c("am"),]
#generating the dataset with EcM only
exudate_EcM <- exudate[exudate$myco %in% c("em"),]
exudate_AM$er_mass_log <- log(exudate_AM$er_mass)
exudate_EcM$er_mass_log <- log(exudate_EcM$er_mass)


p12 <- ggplot(exudate, aes(x=reorder(species, er_mass_log), y=er_mass_log, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~phylo) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~g^-1)))+
  ggtitle("Mass-specific exudation rate among groups: AM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=90))

p12 + facet_wrap(~phylo)

p13 <- ggplot(exudate, aes(x=reorder(species, er_mass_log), y=er_mass_log, fill=species)) + 
  geom_boxplot() +
  facet_wrap(~leaf) + stat_compare_means(method = "anova", size = 2.5)+
  xlab("Species with MF Type") + ylab(expression(log~Exudation~rate~ ~(ugC~d^-1~g^-1)))+
  ggtitle("Mass-specific exudation rate among groups: AM Myco. Type") + theme(axis.text.x = element_text(size = 10, angle=90))

p13 + facet_wrap(~leaf)

```



**Statisitical Analyses**
```{r, echo=FALSE, eval=FALSE}
#scaled by fine root SA
hist(exudate$er_sa_fine_log)
qqnorm(exudate$er_sa_fine_log)
qqline(exudate$er_sa_fine_log)
```

```{r, echo=FALSE, eval=TRUE}
#Shapiro-Wilk test (p>0.05 = normal)
shapiro.test(exudate$er_sa_fine_log)
#the data is not normal

#ANOVA
ersa.aov<-aov(er_sa_log ~ species, exudate)
ersa.aov
qqnorm(residuals(ersa.aov))
qqline(residuals(ersa.aov))

shapiro.test(residuals(ersa.aov))
#get output
summary(ersa.aov)
mod_means_contr <- emmeans::emmeans(object = ersa.aov,
                                    pairwise ~ "species",
                                    adjust = "tukey")
mod_means <- multcomp::cld(object = mod_means_contr$emmeans,
                           Letters = letters)
mod_means
```

According to the tukey test, the only difference in exudation rate was between larch and the lowest 4 trees (pawpaw, bald cypress, hickory, and sawara cypress)

# How did exudation & N uptake differ between myco type, phylo, and leaf habit?

**First, we look at the normality of data nested with respect to plant functional groups (i.e., myco, phylo, and leaf types).

```{r, echo=FALSE, eval=TRUE}
hist(exudate$er_sa_fine_log)
qqnorm(exudate$er_sa_fine_log)
qqline(exudate$er_sa_fine_log)

#t.test, assuming log-transformation permits robustness

am<-subset(exudate$er_mass_log, exudate$myco=="am")
em<-subset(exudate$er_mass_log, exudate$myco=="em")

t.test(er_mass_log ~ myco, data=exudate, alternative=("two.sided"))

qqnorm(am, datax=TRUE)
qqline(am, datax=TRUE)
qqnorm(em, datax=TRUE)
qqline(em, datax=TRUE)

```

Welch's test does not only shows no significance in exudation (log-transformed) between mycorrhizal types (p-value=0.135), but exudation is not normally distributed.  

**ANOVA using linear mixed model**

ANalysis Of VAriance using linear mixed model is robust in that we compare means across more than two populations despite unequal sample sizes (or unbalanced design) and that the data contains continuous response variables and factor variables that defines the groups. 

```{r, echo=TRUE, eval=TRUE}
library(lme4)
lme <- lmer(data = exudate, er_mass_log ~ myco + phylo + myco:phylo + (1|species))
summary(lme)
lsmeans(lme, pairwise~myco*phylo, adjust="tukey")

```

No significant difference between myc types or phylo groups, but we constructed a linear mixed model to look at the interaction of myco and phylo. There was a significant difference between EM gymno and EM angio (p < .05). Marginally significant difference between EM gymno and AM gymno (.076).

**Mass-specific exudation rate among plant function groups**

```{r}
ggplot(exudate, aes(x = myco, y = er_mass_log, fill = phylo)) +
  geom_boxplot() +
  labs(x = "Mycorrhizal Type",
       y = expression(atop("log Exudation Rate", paste((ugC~d^-1~g^-1))))) +
  scale_x_discrete (limits = c("am", "em"),
                    labels = c("AM", "ECM")) + stat_compare_means(method = "anova", size = 3.5) +
  theme_linedraw() +
  geom_boxplot(notch=F,
               lwd=.5,
               stat="boxplot",
               outlier.shape = NA) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill=NA, colour = "black", size=.7),
        axis.title.x = element_text(margin = margin(t = 10, b=5), size=20),
        axis.title.y = element_text(margin = margin(l = 5, r=8), size=20),
        axis.text.x= element_text(margin = margin(t = 10), size=12),
        axis.text.y=element_text(margin = margin(r = 10), size=12),
        legend.title = element_text(colour="black", size=16),
        legend.text = element_text(size=14)) +
  scale_fill_manual(name = "Phylogenetic Group",
                    limits = c("angio", "gymno"),
                    labels = c("Angiosperm", "Gymnosperm"),
                    values = c("#CFB436", "#438029")) +
  ggtitle("Mass-specific exudation rate among groups: Myco.+Phylo Type") + theme(axis.text.x = element_text(size = 10, angle=0)) +
  expand_limits(y = 9.25)
```

EcM gymnosperm (larch, spruce, hemlock) and EcM angiosperm (hickory, SWO) show a significant difference in exudation rate, while in AM trees there is no difference between phylogenetic group. We should look into the patterns of not having much differences with respect to plant function types may show different above- and belowground strategies.   

**Net N uptake rate among plant function groups**
```{r}
ggplot(exudate, aes(x = myco, y = N_uptake, fill = phylo)) +
  geom_boxplot() +
  labs(x = "Mycorrhizal Type",
       y = expression(atop("log net N uptake Rate", paste((mgN~root~g^-1~d^-1))))) +
  scale_x_discrete (limits = c("am", "em"),
                    labels = c("AM", "ECM")) + stat_compare_means(method = "anova", size = 3.5) +
  theme_linedraw() +
  geom_boxplot(notch=F,
               lwd=.5,
               stat="boxplot",
               outlier.shape = NA) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_rect(fill=NA, colour = "black", size=.7),
        axis.title.x = element_text(margin = margin(t = 10, b=5), size=20),
        axis.title.y = element_text(margin = margin(l = 5, r=8), size=20),
        axis.text.x= element_text(margin = margin(t = 10), size=12),
        axis.text.y=element_text(margin = margin(r = 10), size=12),
        legend.title = element_text(colour="black", size=16),
        legend.text = element_text(size=14)) +
  scale_fill_manual(name = "Phylogenetic Group",
                    limits = c("angio", "gymno"),
                    labels = c("Angiosperm", "Gymnosperm"),
                    values = c("#CFB436", "#438029")) +
  ggtitle("Net N uptake rate among groups: Myco.+Phylo Type") + theme(axis.text.x = element_text(size = 10, angle=0)) +
  expand_limits(y = 9.25)
```

## 2) Does root exudation rate vary with plant or root traits? Which variables best predict root exudation rates?

To answer the questions, we need to construct mixed linear models that are similar to ANOVA and regressions but take plot-level replicates into account. To be continued.

First, we generate a correlation matrix plot (called a Correlogram); Positive correlations are displayed in a blue scale while negative correlations are displayed in a red scale.

```{r, echo=FALSE, eval=TRUE}
# More visualization with bivariate analysis

#exudate$srl_log<- log(exudate$srl) 
#exudate$rtd_log<- log(exudate$rtd)
#exudate$X.N_log<- log(exudate$X.N)
#exudate$C.N_log<- log(exudate$C.N)
#exudate$mass_g_log<- log(exudate$mass_g)
#exudate$tip_mass_g_log<- log(exudate$tip_mass_g)
#exudate$diameter_log<- log(exudate$diameter)
#exudate$volume_log<- log(exudate$volume)
#exudate$sa_fine_log<- log(exudate$sa_fine)
#exudate$N_uptake_log<- log(exudate$N_uptake) 
#exudate$total_exudate_log <- log(exudate$total_exudate) 

library(GGally)
library(ggplot2)
library(Hmisc)
setwd("/Users/young/Desktop/Git_Phillips_Lab/PhillipsLab/Morton_analyses/")
exudate <- read.csv("final exudate data_master.csv")

exudate$total_exudate <- exudate$total_exudate_μg_day
exudate$mass <- exudate$mass_g
exudate$tip_mass <- exudate$tip_mass_g
exudate$length <- exudate$length_cm
exudate$er_mass_tip <- as.numeric(exudate$er_mass_tip)
exudate$n <- exudate$X.N
exudate$c <- exudate$X.C
exudate$c_n <- exudate$C.N
exudate$total_exudate_log <- log(exudate$total_exudate) 
exudate$N_uptake_log<- log(exudate$N_uptake) 


ggpairs(exudate, columns = c("total_exudate_log","N_uptake_log", "srl", "rtd","sra","length","diameter","sa_fine", "n","mass_g","branching_intensity", "volume"), title = "Bivariate analysis of Exudation rates vs. root traits (alpha=0.05)", upper = list(continuous = wrap("cor",
        size = 3)),
    lower = list(continuous = wrap("smooth",
        alpha = 0.5,
        size = 0.1)),
    mapping = aes(color = myco))

```

Based on the bivariate analysis of total exudation rates (log-transformed) vs. root traits, we examine more closely on the relationship of exudation rate with the following fine root traits; net N uptake rate, RTD, fineroot surface area, biomass, and branching intensity which exhibit statistical significance. 

The pearson's correlation matrix shows that exudation rates in these samples are correlated with net N uptake rate, RTD, fine root surface area, biomass, and branching intensity, while the degree of correlation differ among myco types. Especially, exudation rate of EM species can be explained by greater number of root morphological traits such as length, Diam, RTD, Surf_area, Root_vol, and volume. The surface area demonstrates the strongest relationship with exudation rate, and other traits that share root area in their measurements also show strong correlation. We need to investigate if exudation rates of EM species are really more correlated with root morphology and why. 

## In-depth correlation analyses

We built to investigate the correlation of specific root traits that appear to have significant relationship with exudation rate.

* Note: all the correlation coefficient displayed in the figures below are the coefficient of determination (R squared, not squared Pearson correlation coefficient (r squared)). 

# Does root tissue density (RTD) corelate with exudate production and the correlation varies among plant functional groups?

# Exudation rate vs. RTD

```{r, echo=FALSE, eval=TRUE}
#adding group means for each species to the plot
library(ggpmisc)
library(dplyr) 
#taking group mean

gd <- exudate %>%  group_by(species) %>% summarise(rtd = mean(rtd),
                  total_exudate_log  = mean(total_exudate_log))
gd

plot_RTD_spp_mean <- ggplot(exudate, aes(x = rtd, y = total_exudate_log)) +
geom_point(alpha = .5, aes(color=species, shape=leaf)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd, size = 4, aes(color=species)) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("Species")) + labs(title = "Relationship of root exudation with RTD") + xlab(expression(RTD~(g~cm^-3))) + ylab(expression(Total~log~Exudate~ ~(ugC~day^-1)))
#geom_smooth(method = "lm")
plot_RTD_spp_mean + geom_smooth(method = "lm")
```
Among 11 species we can't find a significant relationship between RTD and exudation rate. We nested values to mycorrhizal groups to see if the strength of relationship vary among myco. groups.

# Exudation rate vs. RTD nested in diff. functional groups

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(exudate,aes(x = rtd, y = total_exudate_log)) + 
  geom_point(alpha = .75, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("phylo")) + labs(title = "Relationship of root exudation with RTD (AM vs. EcM)") + xlab(expression(RTD~(g~cm^-3))) + ylab(expression(Exudation~rate~ ~(ugC~day^-1)))
#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~myco, scales = "free") + geom_smooth(method = "lm") 
```

The relationship of exudation rate with RTD nested in myco types exhibits that only EcM trees have a significant negative correlation (R-squared = 0.35, p-value <0.001). The figures exhibit the stronger correlation between exudation rate and RTD in EcM trees. RTD is relevant to fast root metabolism (termed as explitative strategy), while it is independent of effective morphological trait such as SRL (Sun et al. 2020; Wen et al. 2021). Since faster root exudation is thought to be related to low RTD (low root construction cost), to see if the tight correlation of exudation rate with root metabolic traits (root[N] and respiration) and RTD as suggested by Sun et al. (2020), we need to examine root[N] more closely. This needs further cross examination of the correlation between exudation rate and SRL suggested by Wang et al. (2019) and Meier et al. (2019). 

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(exudate,aes(x = rtd, y = total_exudate_log)) + 
  geom_point(alpha = .75, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("phylo")) + labs(title = "Relationship of root exudation with RTD (AM vs. EcM)") + xlab(expression(RTD~(g~cm^-3))) + ylab(expression(Exudation~rate~ ~(ugC~day^-1)))
#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~phylo, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~leaf, scales = "free") + geom_smooth(method = "lm") 
```

As for other functional groups (i.e., phylo and leaf types), exudation rate show no relationship with RTD.

# Does net N uptake corelate with exudate production?

```{r, echo=FALSE, eval=TRUE}
#adding group means for each species to the plot
#taking group mean
gd <- exudate %>%  group_by(species) %>% summarise(N_uptake_log = mean(N_uptake_log),
                  total_exudate_log  = mean(total_exudate_log))
gd

plot_RTD_spp_mean <- ggplot(exudate, aes(x = N_uptake_log, y = total_exudate_log)) +
geom_point(alpha = .5, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd, size = 4, aes(color=species)) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("Species")) + labs(title = "Relationship of root exudation with Net N uptake rate") + xlab(expression(Net~N~ uptake~rate~(mgN~root~d^-1))) + ylab(expression(Total~log~Exudate~ ~(ugC~day^-1)))
#geom_smooth(method = "lm")
plot_RTD_spp_mean + geom_smooth(method = "lm")

```

It needs some interpretation....


```{r, echo=FALSE, eval=FALSE}
gd2 <- exudate %>%  group_by(species) %>% summarise(N_uptake_log = mean(N_uptake_log),
                  total_exudate_log  = mean(total_exudate_log))
gd2

plot_N_uptake_spp_mean <- ggplot(exudate, aes(x = N_uptake_log, y = total_exudate_log)) +
geom_point(alpha = .5, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd2, size = 4, aes(color=species)) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("phylo")) + labs(title = "Relationship of root exudation with Net N uptake rate") + xlab(expression(Net~N~ uptake~rate~(mgN~root~d^-1))) + ylab(expression(Total~log~Exudate~ ~(ugC~day^-1)))
#geom_smooth(method = "lm")
plot_N_uptake_spp_mean + geom_smooth(method = "lm")

```

# Do the relatioinships vary among diff. functional groups?

```{r, echo=FALSE, eval=TRUE}
library(ggpmisc)

plt <- ggplot(exudate,aes(x = N_uptake_log, y = total_exudate_log)) + 
  geom_point(alpha = .75, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("phylo")) + labs(title = "Relationship of root exudation with Net N uptake rate (AM vs. EcM)") + xlab(expression(Net~N~ uptake~rate~(mgN~root~d^-1))) + ylab(expression(Exudation~rate~ ~(ugC~day^-1)))
#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~myco, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~phylo, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~leaf, scales = "free") + geom_smooth(method = "lm") 
```

The correlation between exudation and net N uptake rate for both myco. and phylo. types show positive correlation while the strength of relationships varies among plant functional types. Need to know what the main drivers would be to explain the relationships.

# Does amount of root corelate with carbon production?

```{r, echo=FALSE, eval=TRUE}
#adding group means for each species to the plot

#taking group mean

gd <- exudate %>%  group_by(species) %>% summarise(mass_g = mean(mass_g),
                  total_exudate_log  = mean(total_exudate_log))
gd

plot_mass_g_spp_mean <- ggplot(exudate, aes(x = mass_g, y = total_exudate_log)) +
geom_point(alpha = .5, aes(color=species, shape=leaf)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd, size = 4, aes(color=species)) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("Species")) + labs(title = "Relationship of root exudation with root mass") + xlab(expression(Root~mass~(g))) + ylab(expression(Total~log~Exudate~ ~(ugC~day^-1)))
#geom_smooth(method = "lm")
plot_mass_g_spp_mean + geom_smooth(method = "lm")
```

This positive strong correlation (R-squared = 0.28, p-value <0.001) represents that the more mass in a cuvette leading to higher exudation, showing our measurements were working well.  


# Does the strength of correlation vary among functional groups?
```{r, echo=FALSE, eval=TRUE}
plt <- ggplot(exudate,aes(x = mass_g, y = total_exudate_log)) + 
  geom_point(alpha = .75, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("phylo")) + labs(title = "Relationship of root exudation with root biomass (myco, phylo, leaf types)") + xlab(expression(Root~mass~(g))) + ylab(expression(Exudation~rate~ ~(ugC~day^-1)))
#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~myco, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~phylo, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~leaf, scales = "free") + geom_smooth(method = "lm") 

```

The relationship of root exudation with net N uptake rate exhibits different degrees of correlation in 3 different plant functional groups (myco, phylo, and leaf types). Generally, in EcM and gymnosperm we can find stronger relationships.

# Other important root traits that show insignificant correlation; srl and sra

** Does net SRL correlate with exudate production?

```{r, echo=FALSE, eval=TRUE}
#adding group means for each species to the plot

#taking group mean

gd <- exudate %>%  group_by(species) %>% summarise(srl = mean(srl),
                  total_exudate_log  = mean(total_exudate_log))
gd

plot_mass_g_spp_mean <- ggplot(exudate, aes(x = srl, y = total_exudate_log)) +
geom_point(alpha = .5, aes(color=species, shape=myco)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd, size = 4, aes(color=species)) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("Species")) + labs(title = "Relationship of root exudation with specific root length (SRL)") + xlab(expression(SRL~(cm^2~g^-1))) + ylab(expression(Total~log~Exudate~ ~(ugC~day^-1)))
#geom_smooth(method = "lm")
plot_mass_g_spp_mean + geom_smooth(method = "lm")
```

Clearly, our measurements did not show any significant correlation between exudation and specific root length. Thus, we examine if root length alone may correlate to exudation rate. 

# Other important root traits that show insignificant correlation; root length vs. exudation

```{r, echo=FALSE, eval=TRUE}
#adding group means for each species to the plot

#taking group mean

gd <- exudate %>%  group_by(species) %>% summarise(length = mean(length),
                  total_exudate_log  = mean(total_exudate_log))
gd

plot_mass_g_spp_mean <- ggplot(exudate, aes(x = length, y = total_exudate_log)) +
geom_point(alpha = .5, aes(color=species, shape=myco)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, label.x.npc = "right", vstep = 0.05) +  geom_point(data = gd, size = 4, aes(color=species)) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("Species")) + labs(title = "Relationship of root exudation with root length") + xlab(expression(Root~length~(cm))) + ylab(expression(Total~log~Exudate~ ~(ugC~day^-1)))
#geom_smooth(method = "lm")
plot_mass_g_spp_mean + geom_smooth(method = "lm")
```

# Does the strength of correlation between  vary among functional groups?

```{r, echo=FALSE, eval=TRUE}
plt <- ggplot(exudate,aes(x = length, y = total_exudate_log)) + 
  geom_point(alpha = .75, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("phylo")) + labs(title = "Relationship of root exudation with root biomass (myco, phylo, leaf types)") + xlab(expression(Root~length~(cm))) + ylab(expression(Exudation~rate~ ~(ugC~day^-1)))
#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~myco, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~phylo, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~leaf, scales = "free") + geom_smooth(method = "lm") 

```

The notable difference in the correlation of root exudation with length and SRL is not only interesting but it also signifies the importance in the potential of different morphologies driving differential effects on the rhizosphere. We need to decipher the implication more.

Given the high correlation between root mass and length, more root segments lead to more exudation, representing the important roles of RTD that can be translated into root consruction cost and different foraging strategies. 

# Other important root traits that show insignificant correlation; branching intensity vs. exudation

```{r, echo=FALSE, eval=TRUE}
plt <- ggplot(exudate,aes(x = branching_intensity, y = total_exudate_log)) + 
  geom_point(alpha = .75, aes(color=species, shape=phylo)) + stat_poly_eq(formula = y ~ x, aes(label = paste(..eq.label.., ..rr.label.., ..p.value.label.., sep = "*`,`~")), parse = TRUE, size = 2.4, label.x.npc = "right", vstep = 0.05) +  theme_bw() + guides(color = guide_legend("species"),  shape = guide_legend("phylo")) + labs(title = "Relationship of root exudation with branching intensity (myco, phylo, leaf types)") + xlab(expression(Branching~intensity)) + ylab(expression(Exudation~rate~ ~(ugC~day^-1)))
#another way to print stat. info.  
#stat_regline_equation(label.y = 400, aes(label = ..eq.label..)) +
  #stat_regline_equation(label.y = 350, aes(label = ..rr.label..)) + stat_fit_glance(method = 'lm',
                  #method.args = list(formula = "y~x"), geom = 'text', aes(label.y = 300, label = paste("P = ", #signif(..p.value.., digits = 4), sep = "")))

plt + facet_wrap(~myco, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~phylo, scales = "free") + geom_smooth(method = "lm") 
plt + facet_wrap(~leaf, scales = "free") + geom_smooth(method = "lm") 

```

In the relationship between exudation and branching intensity, only EcM and gymnosperm group shows a significant relationships. We need to know why and what effects branching intensity may have on euxdation rate. 


```{r, echo=FALSE, eval=TRUE}
library(permute)
library(lattice)
library(vegan)

PCA_rtrait <- read.csv("Exudation_Collection_compiled_072822.csv")

#head(PCA_rtrait)

rtrait_new <- subset (PCA_rtrait, select = -c(SampleID,Myco,Phylo,Leaf,Tip_Mass))
rda_PCA_rtrait <- rda(rtrait_new[,-1], scale=TRUE)
#plot(rda_PCA_rtrait)
biplot(rda_PCA_rtrait, type = c("text", "points")) 
#summary(rda_PCA_rtrait, display=NULL)
#screeplot(rda_PCA_rtrait, main=NULL)
#screeplot(rda_PCA_rtrait, main=NULL, bstick=T, type='l')

#An example of cluster analysis 
#plot(hclust(dist(scale(rtrait[,-c(1:5)]))))

#An example of PerMANOVA (Permutational ANOVA) analysis 
#adonis(vegdist(rtrait[,6:14]) ~ Myco*Species, data=rtrait)

#res.man <- manova(cbind(Exudation_rate, RTD) ~ Species, data = rtrait)
#summary(res.man)

```
I need a special lesson how to consruct PCA plots and what the whole procedure means.

